### 2022-03-11 Report/NO2.Rmd
Upload New File

---
title: "NO2  case study (5 stations)"
author: "HW"
date: "March 12, 2019"
output: html_document
---

Data set from the paper  by Cooley,  Davis, Naveau (The Annals of Applied Statistics 2012, Vol. 6, No. 4, 1406–1429). 
 Data available  at  http://www.stat.colostate.edu/~cooleyd/DataAndCode/PredExtremes/.

## Description

- Variable:  daily maximum hourly NO2 measurement.

- Five stations located in Washington DC and nearby Virginia: Alexandria (alx), McMillan (mc), River Terrace (rt), Takoma School (ts) and Arlington (arl).

- Period: 5000 daily NO2 measurements recorded between 1995 and 2010, keeping only those records for which all five stations have measurements, resulting in 4497 daily measurements.

- Rounding: because the data were truncated to the nearest ppb, a uniform random variable on the interval [−0.5, 0.5] was added to the data so that they behave more like the underlying continuous variable.

```{r}

## Load data and display first and last 5 lines

load("nitrogenDioxideMsmts.RData")
head(no2Msmts); tail(no2Msmts)
```

```{r}

## Load graphical package lattice 
library(lattice)

## Times series at each of the 5 stations
dates=as.Date(no2Msmts[,1], origin = "1970-01-01")
xyplot(no2Msmts[,2]~ dates,pch=19,ylab="NO2 at Alexandria")
xyplot(no2Msmts[,3]~ dates,pch=19,ylab="NO2 at Arlington")
xyplot(no2Msmts[,4]~ dates,pch=19,ylab="NO2 at McMillan ")
xyplot(no2Msmts[,5]~ dates,pch=19,ylab="NO2 at River Terrace")
xyplot(no2Msmts[,6]~ dates,pch=19,ylab="NO2 at Takoma School")
```

## Examine dependence on weekdays

On Saturday and Sunday there is less NO2 for all series.

```{r}
## data object is a matrix, create a data frame
class(no2Msmts)
NO2=as.data.frame(no2Msmts)
names(NO2)
```

```{r}

## in case the locale is not in english
Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

```{r}

## dependence on weekdays: on Sat and Sun there is decrease of NO2
day= factor( weekdays(dates,abbreviate=TRUE), levels=c("Mon","Tue","Wed","Thu", "Fri", "Sat", "Sun"),ordered=TRUE)

bwplot(day~NO2$ts, main="Tacoma school station",xlab="NO2")

```

## Examine dependence on month

Strangely the median is highest for Feb, Mar, Apr for all series

```{r}
month =factor( months(dates,abbreviate=TRUE),levels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE)

bwplot(month~NO2$alx);bwplot(month~NO2$arl);bwplot(month~NO2$mc);bwplot(month~NO2$rt);bwplot(month~NO2$ts)

```

## QQ plots

Empirical quantiles vs normal quantiles

```{r}
qqmath(NO2$alx[month=="Jul"])
```

Empirical quantiles vs exponential quantiles

```{r}
qqmath(NO2$alx[month=="Jul"],distri=qexp)
```

## Extreme value distribution fit

We use the library ismev that goes with the book by Stuart Coles

```{r}
# the library ismev needs to be downloaded from CRAN to your machine
# install.packages("ismev")

## Load library ismev
library(ismev)
```

### GEV modelling for each station

Modelling all data without distinguishing days and months, the tail is not fitted well at all for each of the series.

```{r]
## all days and months together
gev.diag(gev.fit(NO2$alx));gev.diag(gev.fit(NO2$arl));gev.diag(gev.fit(NO2$mc));gev.diag(gev.fit(NO2$rt));gev.diag(gev.fit(NO2$ts))
```

### GEV modelling performed for each month

Better fits are obtained for individual months

```{r}
## for a specific month
gev.diag(gev.fit(NO2$alx[month=="Jul"]));gev.diag(gev.fit(NO2$arl[month=="Jul"]));gev.diag(gev.fit(NO2$mc[month=="Jul"]));gev.diag(gev.fit(NO2$rt[month=="Jul"]));gev.diag(gev.fit(NO2$ts[month=="Jul"]))

```

## Mean excess plot

```{r}
mrl.plot(NO2$alx[month=="Jul"])
```

## Scatter plot matrix for the 5 stations

No distinction of months.

Observe the extremes: are they common for a pair of stations ?

```{r}
splom(NO2[,2:6],pch=19)

```

## Scatter plot matrix for the 5 stations

Using a different colour for each month.

```{r}
splom(NO2[,2:6],group=as.factor(month),auto.key=TRUE)
```