### 2022-03-16 Practicals/Day1_MEUSE.Rmd
Saved changes to: Code_examples/plot_wind.R Practicals/Day1_MEUSE.Rmd Practicals/Day1_MEUSE.html

---
title: "Meuse river annual maximal discharge"
author: "HW"
date: "May 24, 2018"
output: html_document
---

Meuse annual maximal discharge
==

```{r, echo=FALSE}
# LIBRARY: Intro Stat Modeling Extreme Values
library(ismev)
```

```{r}
# import meuse data
source("Day1_meuse.data.R")
attach(meuse)
require(lattice)
names(meuse)
xyplot(Discharge~Year,pch=5)
```

Meuse annual maximal discharge
==

```{r}

dotplot(Discharge)
bwplot(Discharge)
histogram(Discharge,n=20)
```

Normal and exponential QQ plots
==

```{r}
qqmath(Discharge,xlab="Normal quantiles")
#abline(a = 1, b = 0)
# QQplot for Extreme Value Analysis
qqmath(Discharge,distribution = qexp,xlab="Exponential quantiles")

```

Two versions of the excess plot
==

```{r}
# evir library: excess plot
library(evir)
meplot(Discharge)
# ismev library: excess plot
mrl.plot(Discharge)
# notice that meplot has a parameter omit=3: three largest values are omitted
?meplot

```

GEV and Gumbel model fits
=

```{r}
meuse.GEV=gev.fit(Discharge)
meuse.GUM=gum.fit(Discharge)
# plot diagnostics
gev.diag(meuse.GEV)
gum.diag(meuse.GUM)
```
```{r}
gev.profxi(meuse.GEV,-.3,.3)
gev.prof(meuse.GEV,100, 2500,4500)
```
CONCLUSION
=

Same conclusion as Beirlant et al.:
Gumbel model is better than GEV model.

NEW ANALYSIS OF THAT DATA
==

Discussion: are there outliers?

```{r}
plot(meuse)
lines(lowess(meuse),col=2)

```

Which years?
==

```{r}
# there seems to be an outlier: which year?
plot(meuse)
###  For interactivity execute the commands of chunk in the CONSOLE
# CLICK on point in Plot window [ESC  to end]
identify(meuse)
which.max(meuse$Discharge)
```

Outliers
==

```{r}
Year[15]; Year[83]

```
Remove outliers
==

```{r}
mm2 = data.frame(cbind(Year[c(-15,-83)],Discharge[c(-15,-83)]))
colnames(mm2) = c("Year","Discharge")
```

Recompute and display the overall trend without outlier
==

```{r}
plot(meuse) # with outlier
# using "loess"" instead of "lowess""
lines(meuse$Year,
predict(loess(Discharge ~ Year,
data=meuse,span=1.2)) ,col=4)

plot(mm2) # without outlier
lines(mm2$Year,
predict(loess(Discharge ~ Year,
data=mm2,span=1.2)) ,col=2)

```

Including the trend into the GEV analysis
==

```{r}
## storing the trend
meuse.loess =
as.data.frame(
predict(loess(Discharge ~ Year,
data=mm2,span=1.2)))
colnames(meuse.loess)="CE"

## storing the analysis

meuse.GEVloess =
gev.fit(mm2$Discharge,ydat=meuse.loess,mul=1)
meuse.GUMloess =
gum.fit(mm2$Discharge,ydat=meuse.loess,mul=1)

# plot diagnostics
gev.diag(meuse.GEVloess)
gum.diag(meuse.GUMloess)

```
```{r}
## compare the negative log-likelihoods (nllh)
meuse.GEVloess$nllh
meuse.GUM$nllh
```

CONCLUSION
==

GEV model fits the residuals better than Gumbel model.