### 2022-03-19 Datasets/Draft#1.Rmd
draft1

---
title: "Extreme Value Statistics Draft 1"
Dataset: Wind 
Group: 1
Name: Chaitanya Arora, Diandra Gramma 
output: html_document
---

## Introduction ##

Before deep diving into the analysis and coding portion,we should analyse the type of data that we have and what does it mean.

The wind datasets is contains 2 files:

1. "coordinates": This file contains the coordinates of the 5 places in western Europe which we will analyse during the process of the project

2. "dataset": this is a large matrix containing maximum speed in the duration of 3 hours which lasted for more than 3sec. 
The data is from 10th January 1979 - 4th Jan 2014 (35 years)

## Step 1: Basic Analysis of the data ##

1.1 The graphical representation of the 5 places that we are dealing with

```{r}

load("Datasets/1_Wind.RData")
library(ggmap)

bbox <- c(-21.375,  40.875,  16.875,  61.125)

map <- get_map(location = bbox, source="stamen", maptype = "toner" , color="bw")

dataToPlot = data.frame(X = coordinates[, 1], Y = coordinates[, 2])
mapPoints <- ggmap(map) +
  geom_point(
    data = dataToPlot,
    aes(x = X, y = Y),
    color = "red",
    alpha = 1,
    size = 3,
    shape = 16,
  ) +
  labs(title = "", fill = "") +
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text=element_text(size=14))
print(mapPoints)

```